### 2014-11-09 tex/proofs.tex
Locked math. Switch to paper

%we need to define how r' is arrived at for this one to work...
\begin{lemma}[\textrm{F{\footnotesize IELD}} \textrm{A{\footnotesize CCESS}} preserves $\sqsubset\ \subseteq F_\sim(\sqsubset)$]
If two field access states are related in the represented by relation,
then they are also related in the functional associated to bisumlation.
\label{lem:access}
$$
\forall p \in S_\mathit{FA}\ (\forall q \in S_\mathit{FA}\ (p \sqsubset q \Rightarrow (p\ q) \in F_\sim(\sqsubset)))
$$
\end{lemma}

\begin{proof}
Proof by contradiction: assume $p \sqsubset q \wedge (p\ q) \not\in F_\sim(\sqsubset)$.

The case where neither $p$ nor $q$ have successors is trivially satisfied by
\defref{bisimulation} since the conditions for inclusion in
$F_\sim(\sqsubset)$ are vacuously met. That is a contradiction.

Consider now the case where $p$ and $q$ have successor states. The
statement $p \sqsubset q$ (\defref{representation}) means that $p$ and
$q$ are only differentiated by their heaps and global constraints
since the environment ($\eta$), expression ($e$), and continuation
($k$) are the same in both states.

Ignoring the heaps, and given that $p$ and $q$ currently have
the same environment, expression, and continuation, all successors
of $p$ related by \textrm{F{\footnotesize IELD}}
\textrm{A{\footnotesize CCESS}} in $\rgse$ (\figref{fig:lazy}) have
the same unchanged environment from $p$, $\eta$, new expression
$r^\prime = \mathrm{stack}_\cfgnt{r}()$, and continuation $\cfgnt{k}$
from the old continuation in $p$ having completed the field
access. Similarly, the one successor of $q$ related by $\rsym$
(\figref{fig:symfield}), has that same $\eta$, $r^\prime =
\mathrm{stack}_\cfgnt{r}()$, and $\cfgnt{k}$. As such, every successor
of $p$/$q$ has a matching successor of $q$/$p$ that agrees on the
environment, expression, and continuation meeting the first condition
necessary to relate the successors in $\sqsubset$.

Turning now to the heaps, \textrm{F{\footnotesize IELD}}
\textrm{A{\footnotesize CCESS}}, in $\rgse$ and $\rsym$, initializes
uninitialized locations in the heap on reference $\cfgnt{r}$ for field
$\cfgnt{f}$. The heaps in $p$ and $q$ are still homomorphic after
initialization, and the heap constraint in the homomorphism is still
valid by \lemref{lem:init}. Let
$(\cfgnt{L}^\prime_p\ \cfgnt{R}^\prime_p) \rightharpoonup_h
(\cfgnt{L}^\prime_q\ \cfgnt{R}^\prime_q)$ be those new heaps and
the homomorphism after initialization on the heap in $p$ with
$\rightarrow_I^*$ (\figref{fig:lazyInit}) and the heap in $q$ with
$\rightarrow_S^*$ (\figref{fig:symInit}).

After initialization, \textrm{F{\footnotesize IELD}} \textrm{A{\footnotesize CCESS}} for $\rgse$ creates a new heap
$(\cfgnt{L}^\prime_p [\cfgnt{r}^\prime \mapsto \lp\phi^\prime\ l^\prime\rp]\ \cfgnt{R}^\prime_p$

and for $\rsym$ it relates the state 
$$ 
s_\sym^\prime = \lp \cfgnt{L}_{s}[\cfgnt{r}^\prime \mapsto \mathbb{VS}\lp \cfgnt{L}_{s},\cfgnt{R}_{s},\cfgnt{r},\cfgnt{f},\phi_g\rp ]\ \cfgnt{R}_{s}\ \phi_g\ \eta\ \cfgnt{r}^\prime\ \cfgnt{k}\rp 
$$

The functional in \defref{} requires and forward
\eqref{eqn:BisimulationForwards} and backward
\eqref{eqn:BisimulationBackwards} relationship between successor
heaps.

\eqref{eqn:BisimulationForwards}
\eqref{eqn:BisimulationBackwards}

We now show that $s_\gse^\prime \sqsubset s_\sym^\prime$. Since $\eta$, $e$, and $k$ are identical between $s_\sym^\prime$ and $s_\gse^\prime $, the first condition is met by default. Now we construct the function $h^\prime$ such that $h^\prime = h$. Observe that since $s_\gse \rightharpoonup_{h} s_\sym$, and since $\cfgnt{R}_\gse$ and $\cfgnt{R}_\sym$ are unchanged from states $s_\gse$ to $s_\gse^\prime$ and $s_\sym$ to $s_\sym^\prime$ respectively, we are guaranteed that $ \cfgnt{r} = \cfgnt{R}_\gse(l,f) \Rightarrow \cfgnt{r} = \cfgnt{R}_\sym(h^\prime(l),f)$. Let $\{(\phi_\gse^\prime\ l^\prime)\} =  \cfgnt{L}_\gse(\cfgnt{R}_\gse(l,f))$. Since $\mathbb{S}(\phi_g \wedge \mathbb{HC}(s_\gse \rightharpoonup_{h} s_\sym))$ is valid, we know that:
 $$(\phi_\sym \wedge \phi_\sym^\prime\ h(l^\prime)) \in \mathbb{VS}\lp \cfgnt{L}_{s},\cfgnt{R}_{s},\cfgnt{r},\cfgnt{f},\phi_g\rp$$ 
From this, we may deduce that:
$$ (\phi_\gse\ l) \in \cfgnt{L}_\gse^\prime(\cfgnt{r}^\prime) \Rightarrow (\phi_\sym \wedge \phi_\sym^\prime\ h^\prime(l))\in \cfgnt{L}_\sym^\prime(\cfgnt{r}^\prime)$$
Since $\cfgnt{r}^\prime$ is the only new addition to $L_\gse^\prime$ and $L_\sym^\prime$, we now know that the assertion above holds for all $l \in \mathcal{L}$. Thus, we have shown that $s_\gse^\prime \rightharpoonup_{h} s_\sym^\prime$. Furthermore, since the constraints in $\mathbb{HC}(s_\gse^\prime \rightharpoonup_{h^\prime} s_\sym^\prime)$ are constructed using conjuncts already present in $ \mathbb{HC}(s_\gse \rightharpoonup_{h} s_\sym)$, we are guaranteed that $\mathbb{HC}(s_\gse^\prime \rightharpoonup_{h^\prime} s_\sym^\prime) \Leftrightarrow \mathbb{HC}(s_\gse \rightharpoonup_{h} s_\sym)$, and therefore $\mathbb{S}(\phi_g \wedge \mathbb{HC}(s_\gse^\prime \rightharpoonup_{h^\prime} s_\sym^\prime))$. This fact, and the fact that $\eta_{\gse} = \eta_{s} ,\ \cfgnt{e}_{\gse} = \cfgnt{e}_{s} ,\ \cfgnt{k}_{\gse} = \cfgnt{k}_{s}$, means that by Definition~\ref{representation} we know $s_\gse^\prime \sqsubset s_\sym^\prime$. We have now shown that:
\begin{equation}
\forall s_\gse^\prime ( s_\gse \rgse s_\gse^\prime \Rightarrow \exists s_\sym^\prime( (s_\sym \rsym s_\sym^\prime )\wedge (s_\gse^\prime\ \sqsubset\ s_\sym^\prime ))  )
\end{equation}

@@

\begin{lemma}[$\rsum^*$ preserves heap determinism]
\label{lem:S-determ}
Given a deterministic heap, $(\cfgnt{L}_0\ \cfgnt{R}_0)$, from a
state with a reference $\cfgnt{r}$ and field $\cfgnt{f}$, the
new heap, $(\cfgnt{L}^\prime\ \cfgnt{R}^\prime)$, from the 
summary machine, $(\cfgnt{L}_0\ \cfgnt{R}_0\ \cfgnt{r}\ \cfgnt{f})
\rsum^*
(\cfgnt{L}^\prime\ \cfgnt{R}^\prime\ \cfgnt{r}\ \cfgnt{f})$, is also deterministic.
\end{lemma}
\begin{proof}
Induction over the number of steps in $\rsum^*$ in \figref{fig:symInit}.

@@

\begin{lemma}[$\rightarrow_\mathit{FA}$ preserves heap determinism]
\label{lem:FA-determ}
Given a state, $s$, with a deterministic heap,
$(\cfgnt{L}\ \cfgnt{R}) =
\mathrm{heap}(s)$, the new heap,
$(\cfgnt{L}^\prime\ \cfgnt{R}^\prime) =
\mathrm{heap}(s^\prime)$, in any state related by the field
access rule, $s \rightarrow_\mathit{FA} s^\prime$,
is also deterministic.
\end{lemma}
\begin{proof}
Proof by definition of $\rightarrow_\mathit{FA}$ in \figref{fig:sym}.

@@

\begin{lemma}[$\rightarrow_\mathit{FW}$ preserves heap determinism]
\label{lem:FW-determ}
Given a state, $s_\sym$, with a deterministic heap,
$(\cfgnt{L}_\sym\ \cfgnt{R}_\sym) =
\mathrm{heap}(s_\sym)$, the new heap,
$(\cfgnt{L}_\sym^\prime\ \cfgnt{R}_\sym^\prime) =
\mathrm{heap}(s_\sym^\prime)$, in any state related by the field
write rule, $s_\sym \rightarrow_\mathit{FW} s_\sym^\prime$,
is also deterministic.
\end{lemma}
\begin{proof}
Proof by definition of $\rightarrow_\mathit{FW}$ in \figref{fig:sym}.

@@

\begin{lemma}[$\rcom$ preserves heap determinism]
\label{lem:J-determ}
Given a state, $s_\sym$, with a deterministic heap,
$(\cfgnt{L}_\sym\ \cfgnt{R}_\sym) =
\mathrm{heap}(s_\sym)$, the new heap,
$(\cfgnt{L}_\sym^\prime\ \cfgnt{R}_\sym^\prime) =
\mathrm{heap}(s_\sym^\prime)$, in any state related by the
Javalite relation, $s_\sym \rcom s_\sym^\prime$, is
also deterministic.
\end{lemma}
\begin{proof}
Proof by definition of $\rcom$ in \figref{fig:javalite-common}.

@@

\begin{theorem}[$\rsym$ preserves heap determinism]
\label{thm:determ}
Given a state, $s_\sym$, with a deterministic heap,
$(\cfgnt{L}_\sym\ \cfgnt{R}_\sym) =
\mathrm{heap}(s_\sym)$, the new heap,
$(\cfgnt{L}_\sym^\prime\ \cfgnt{R}_\sym^\prime) =
\mathrm{heap}(s_\sym^\prime)$, in any state related by the
symbolic relation, $s_\sym \rsym s_\sym^\prime$, is
also deterministic.
\end{theorem}
\begin{proof}
Proof by \lemref{lem:FA-determ}, \lemref{lem:FW-determ}, and \lemref{lem:J-determ}
which represent all the rules that relate states in
$\rsym$.
\end{proof}

>>>>

\begin{comment}
Choose any P and Q such that P is represented by Q on field access. Choose some P' that P transitions too.
Pick some p' in P'. By definition, it is related to a p in P. Use that p. Show how that p is related to p'. Then relate to q'.

Arbitrary P' such that P transitions to P'. Reason over each possible P' (how every many choices might exist--Only one for FA). Pick arbitrary p' in one of the sucessor. Do the forward proof. Show connection in q'. Connect to P' since it holds for arbitrary.

Repeat process for each possible successor P'

For backward, for each q', there must exists a p'. All we need is one p and then that implies there exists a P. Must do for all Q'. 
\end{comment}

\begin{lemma}[\textrm{F{\footnotesize IELD}} \textrm{A{\footnotesize CCESS}} preserves $\sqsubset\ \subseteq F_\sim(\sqsubset)$]
If $P \in 2^{S_\mathit{FA}}$ and $q \in S$ are such that $P \sqsubset q$, then $(P\ q)$ is in the functional associated to bisumlation.
\label{lem:access}
$$
\forall P \in 2^{S_\mathit{FA}}\ (P \sqsubset q \Rightarrow (P\ q) \in F_\sim(\sqsubset))
$$
\end{lemma}

\begin{proof}
Proof by contradiction: assume $P \sqsubset q \wedge (P\ q) \not\in F_\sim(\sqsubset)$.

\lemref{lem:succ} ensures at least one successor of $P$ and $q$. Let $p^\prime \in P^\prime$ be any
state in any successor set and $p \rgse p^\prime$ that state's
predecessor. Let $q^\prime$ be similarly defined relative to $q$.  The states $p$ and $q$, by virtue of the
representation relation, are only differentiated by their heaps and
global constraints since the environment ($\eta$), expression ($r$),
and continuation ($\lp \cfgt{*}\ \cfgt{\$}\ \cfgnt{f} \rightarrow
\cfgnt{k}\rp$) are the same in all states 
(\defref{representation}).

There are two relations defined for field access in both $\rsgse$ and
$\rsym$ for a null and non-null pointing reference (\figref{fig:lazy}
and \figref{fig:fHeap}). In each relation, the environment is
unchanged. For a null pointing reference, in both relations, the
expression is $\cfgt{error}$ and the continuation is $\cfgt{end}$. For
the non-null pointing reference, some care is needed. First each does
use the same continuation $\cfgnt{k}$ (it is the same for $p$ and $q$
by $\sqsubset$. As the environment and continuations are the same,
$r^\prime = \mathrm{stack}_\cfgnt{r}()$, returns the same next largest
number of current stake references from the partition. The first
condition for inclusion in $\sqsubset$, equality between environments,
expressions, and continuations, is met.

Now consider the heaps.  The functional in \defref{bisimulation} requires a forward
\eqref{eqn:BisimulationForwards} and backward
\eqref{eqn:BisimulationBackwards} relationship between successor
heaps. First the forward direction.

\noindent{\textrm{F{\footnotesize IELD}}
\textrm{A{\footnotesize CCESS (NULL)}}}: $p \rightarrow^\ell_\mathit{FA(N)}
p^\prime$ because $l = l_\mathit{null}$ (\figref{fig:lazy}). As $p$
is represented in $q$, $\exists q^\prime\ (q
\rightarrow^\sym_\mathit{FA{N}} q^\prime$ since $l_{\mathit{null}}$
must be feasible on the reference (\figref{fig:fHeap}).

The successors $p^\prime$ and $q^\prime$ are only differentiated by
their heap and global constraint as mentioned previously. Since each heap in each successor is unchanged
from the predecessor state, the same homomorphism from $p \sqsubset q$
establishes that $p^\prime \sqsubset q^\prime$ holds in the forward direction.

\noindent\textrm{F{\footnotesize IELD}}
\textrm{A{\footnotesize CCESS}}: $p \rightarrow^\ell_\mathit{FA}
p^\prime $ because $\cfgnt{l} \neq \cfgnt{l}_\mathit{null}$. As $p$ is represented in $q$ there must be a feasible non-null location: $\exists q^\prime\ (q
\rightarrow^\sym_\mathit{FA} q^\prime)$. The non-null pointing field access relations initialize
uninitialized locations in the heap on reference $\cfgnt{r}$ for field
$\cfgnt{f}$. The heaps in $p$ and $q$ are still homomorphic after
initialization, and the heap constraint in the homomorphism is still
valid by \lemref{lem:init}. Let $(\cfgnt{L}^\prime_p\ \cfgnt{R}^\prime_p)$ and  
$(\cfgnt{L}^\prime_q\ \cfgnt{R}^\prime_q)$ be those new heaps and $\rightharpoonup_h$
the homomorphism after initialization on the heap in $p$ with
$\rightarrow_I^*$ (\figref{fig:lazyInit}) and the heap in $q$ with
$\rightarrow_S^*$ (\figref{fig:symInit}).

The final heap in the successor state $p^\prime$ is
$(\cfgnt{L}^\prime_p [\cfgnt{r}^\prime \mapsto
  \lp\phi^\prime\ l^\prime\rp]\ \cfgnt{R}^\prime_p)$, and for $q^\prime$
the final heap in the successor state is
$(\cfgnt{L}^\prime_q[\cfgnt{r}^\prime \mapsto \mathbb{VS}\lp
  \cfgnt{L}^\prime_q,\cfgnt{R}^\prime_q,\cfgnt{r},\cfgnt{f},\phi^\prime_g\rp
]\ \cfgnt{R}^\prime_q)$; each version adding in the
constraint-location pair or value set respectively on the same
reference $\cfgnt{r}^\prime$. As $r^\prime$ is a stack reference, and
identical in both $p^\prime$ and $q^\prime$, and no new locations are
added to either heap, the existing homomorphism after initialization
still holds: $p^\prime \rightharpoonup_h q^\prime$.

 To relate $p^\prime$ and $q^\prime$ on
$\sqsubset$, it is necessary to create a new homomorphism:
$(\cfgnt{L}^\prime_p\ \cfgnt{R}^\prime_p) \rightharpoonup_{h^\prime}
(\cfgnt{L}^\prime_q\ \cfgnt{R}^\prime_q)$. To that end, construct the
function $h^\prime$ such that $h^\prime = h$.

Observe that since $s_\gse \rightharpoonup_{h} s_\sym$, and since $\cfgnt{R}_\gse$ and $\cfgnt{R}_\sym$ are unchanged from states $s_\gse$ to $s_\gse^\prime$ and $s_\sym$ to $s_\sym^\prime$ respectively, we are guaranteed that $ \cfgnt{r} = \cfgnt{R}_\gse(l,f) \Rightarrow \cfgnt{r} = \cfgnt{R}_\sym(h^\prime(l),f)$. Let $\{(\phi_\gse^\prime\ l^\prime)\} =  \cfgnt{L}_\gse(\cfgnt{R}_\gse(l,f))$. Since $\mathbb{S}(\phi_g \wedge \mathbb{HC}(s_\gse \rightharpoonup_{h} s_\sym))$ is valid, we know that:
 $$(\phi_\sym \wedge \phi_\sym^\prime\ h(l^\prime)) \in \mathbb{VS}\lp \cfgnt{L}_{s},\cfgnt{R}_{s},\cfgnt{r},\cfgnt{f},\phi_g\rp$$ 
From this, we may deduce that:
$$ (\phi_\gse\ l) \in \cfgnt{L}_\gse^\prime(\cfgnt{r}^\prime) \Rightarrow (\phi_\sym \wedge \phi_\sym^\prime\ h^\prime(l))\in \cfgnt{L}_\sym^\prime(\cfgnt{r}^\prime)$$
Since $\cfgnt{r}^\prime$ is the only new addition to $L_\gse^\prime$ and $L_\sym^\prime$, we now know that the assertion above holds for all $l \in \mathcal{L}$. Thus, we have shown that $s_\gse^\prime \rightharpoonup_{h} s_\sym^\prime$. Furthermore, since the constraints in $\mathbb{HC}(s_\gse^\prime \rightharpoonup_{h^\prime} s_\sym^\prime)$ are constructed using conjuncts already present in $ \mathbb{HC}(s_\gse \rightharpoonup_{h} s_\sym)$, we are guaranteed that $\mathbb{HC}(s_\gse^\prime \rightharpoonup_{h^\prime} s_\sym^\prime) \Leftrightarrow \mathbb{HC}(s_\gse \rightharpoonup_{h} s_\sym)$, and therefore $\mathbb{S}(\phi_g \wedge \mathbb{HC}(s_\gse^\prime \rightharpoonup_{h^\prime} s_\sym^\prime))$. This fact, and the fact that $\eta_{\gse} = \eta_{s} ,\ \cfgnt{e}_{\gse} = \cfgnt{e}_{s} ,\ \cfgnt{k}_{\gse} = \cfgnt{k}_{s}$, means that by Definition~\ref{representation} we know $s_\gse^\prime \sqsubset s_\sym^\prime$. We have now shown that:
\begin{equation}
\forall s_\gse^\prime ( s_\gse \rgse s_\gse^\prime \Rightarrow \exists s_\sym^\prime( (s_\sym \rsym s_\sym^\prime )\wedge (s_\gse^\prime\ \sqsubset\ s_\sym^\prime ))  )
\end{equation}

@@

\begin{lemma}[$\rsum^*$ produces well-formed heaps]
\label{lem:S-determ}
Given a deterministic heap, $(\cfgnt{L}_0\ \cfgnt{R}_0)$, from a
state with a reference $\cfgnt{r}$ and field $\cfgnt{f}$, the
new heap, $(\cfgnt{L}^\prime\ \cfgnt{R}^\prime)$, from the 
summary machine, $(\cfgnt{L}_0\ \cfgnt{R}_0\ \cfgnt{r}\ \cfgnt{f})
\rsum^*
(\cfgnt{L}^\prime\ \cfgnt{R}^\prime\ \cfgnt{r}\ \cfgnt{f})$, is also deterministic.
\end{lemma}
\begin{proof}
Induction over the number of steps in $\rsum^*$ in \figref{fig:symInit}.

@@

\begin{lemma}[$\rightarrow_\mathit{FA}$ produces well-formed heaps]
\label{lem:FA-determ}
Given a state, $s$, with a deterministic heap,
$(\cfgnt{L}\ \cfgnt{R}) =
\mathrm{heap}(s)$, the new heap,
$(\cfgnt{L}^\prime\ \cfgnt{R}^\prime) =
\mathrm{heap}(s^\prime)$, in any state related by the field
access rule, $s \rightarrow_\mathit{FA} s^\prime$,
is also deterministic.
\end{lemma}
\begin{proof}
Proof by definition of $\rightarrow_\mathit{FA}$ in \figref{fig:sym}.

@@

\begin{lemma}[$\rightarrow_\mathit{FW}$ produces well-formed heaps]
\label{lem:FW-determ}
Given a state, $s_\sym$, with a deterministic heap,
$(\cfgnt{L}_\sym\ \cfgnt{R}_\sym) =
\mathrm{heap}(s_\sym)$, the new heap,
$(\cfgnt{L}_\sym^\prime\ \cfgnt{R}_\sym^\prime) =
\mathrm{heap}(s_\sym^\prime)$, in any state related by the field
write rule, $s_\sym \rightarrow_\mathit{FW} s_\sym^\prime$,
is also deterministic.
\end{lemma}
\begin{proof}
Proof by definition of $\rightarrow_\mathit{FW}$ in \figref{fig:sym}.

@@

\begin{lemma}[$\rcom$ produces well-formed heaps]
\label{lem:J-determ}
Given a state, $s_\sym$, with a deterministic heap,
$(\cfgnt{L}_\sym\ \cfgnt{R}_\sym) =
\mathrm{heap}(s_\sym)$, the new heap,
$(\cfgnt{L}_\sym^\prime\ \cfgnt{R}_\sym^\prime) =
\mathrm{heap}(s_\sym^\prime)$, in any state related by the
Javalite relation, $s_\sym \rcom s_\sym^\prime$, is
also deterministic.
\end{lemma}
\begin{proof}
Proof by definition of $\rcom$ in \figref{fig:javalite-common}.

@@

\begin{theorem}[$\rsym$ preserves heap well-formedness]
\label{thm:determ}
Given a state, $s_\sym$, with a deterministic heap,
$(\cfgnt{L}_\sym\ \cfgnt{R}_\sym) =
\mathrm{heap}(s_\sym)$, the new heap,
$(\cfgnt{L}_\sym^\prime\ \cfgnt{R}_\sym^\prime) =
\mathrm{heap}(s_\sym^\prime)$, in any state related by the
symbolic relation, $s_\sym \rsym s_\sym^\prime$, is
also deterministic.
\end{theorem}
\begin{proof}
Proof by \lemref{lem:FA-determ}, \lemref{lem:FW-determ}, and \lemref{lem:J-determ}
which represent all the rules that relate states in
$\rsym$.
\end{proof}
